---
title: "GP_pca_and_rf"
author: "Shubh Sharma"
date: "8/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Importing Libraries

```{r}
library(dplyr)
library(ggplot2)
library(reshape2)
library(Information)
library(caTools)
set.seed(10)

# Attach Packages
library(tidyverse)    # data manipulation and visualization
library(kernlab)      # SVM methodology
library(e1071)        # SVM methodology
library(ISLR)         # contains example data set "Khan"
library(RColorBrewer) # customized coloring of plots
```

## Readability and Presentation

```{r}
cp_2 <- c("#FEA47F", "#F97F51")
cp_3 <- c("#2A363B", "#E84A5F", "#FF847C")
cp_5 <- c("#2A363B", "#E84A5F", "#FF847C", "#FECEAB", "#99B898")
cp_8 <- c("#FEA47F", "#F97F51", "#B33771", "#3B3B98", "#58B19F", "#BDC581", "#2C3A47", "#82589F")
```

### Data Loading

```{r}
dem <-read.csv("demogs.csv")
```

### Data Summary

```{r}
# Summary
summary(dem)

str(dem)

# No of rows
nrow(dem)
# 71295


```




### Elimination of Duplicates

```{r}
length(unique(dem$Application.ID)) #demo data set
#71292



#filtering out the IDs that are duplicated

dem %>%
  group_by(Application.ID) %>%
  filter(n() > 1)
#assign new App ID to the duplicate records or drop them





dem <- dem %>%
  group_by(Application.ID) %>%
  filter(n() == 1)



```

### Renaming the columns

```{r}
names(dem)[c(1:12)] <- c("Application_ID", "Age", "Gender", "Marital_Status", "No_Of_Dependents", "Income", "Education", "Profession", "Type_Of_Residence", "Months_In_Current_Residence", "Months_In_Current_Company","Performance_Tag")

```

### Performance Tag

* Eliminated records with no value for Performance_Tag

```{r}
dem$Performance_Tag %>%
  is.na() %>%
  sum()

summary(dem$Performance_Tag)

dem <- dem %>%
  filter(!is.na(Performance_Tag))
```


### Age

```{r}
#=========#
#   Age   #
#=========#

# Check for Age variable rows with NA values
dem$Age %>%
  is.na() %>% sum()
# 0

# Checking for outliers
dem$Age %>%
  quantile(seq(0,1, 0.01))

dem$Age %>%
  boxplot(border = "#6fa058", outcol = "#ee853f")

## Min age is -3
## Some ages are 0
## Capping minimum age to 18
## Since 18 is the minimum age to avail a credit card

dem$Age <- dem$Age %>%
  as.numeric()

dem[(which(dem$Age < 18)), ]$Age <- 18

## Creating age bins
dem$Age %>%
  summary()

# Different Bins
# 1) 16-20
# 2) 21-25
# 3) 26-30
# 4) 31-35
# 5) 36-40
# 6) 41-45
# 7) 46-50
# 8) 51-55
# 9) 56-60
# 10) 61-65

# Age Bins function
age_bin <- function(age=3){
  if(age > 17 && age < 21)
    return ("18-20")
  else if(age > 20 && age < 26)
    return ("21-25")
  else if(age > 25 && age < 31)
    return ("26-30")
  else if(age > 30 && age < 36)
    return ("31-35")
  else if(age > 35 && age < 41)
    return ("36-40")
  else if(age > 40 && age < 46)
    return ("41-45")
  else if(age > 45 && age < 51)
    return ("46-50")
  else if(age > 50 && age < 56)
    return ("51-55")
  else if(age > 55 && age < 61)
    return ("56-50")
  else if(age > 60 && age < 66)
    return ("60-65")
  
}

# Creating Age Bin field
dem$Age_Bin <-  dem$Age %>%
  sapply(age_bin) %>%
  as.factor()

# Plot for Frequency of Age Bins
ggplot(dem, aes(x=Age_Bin, y=..count../1000, fill=Age_Bin)) +
  geom_bar() +
  labs(x="Age Bin", y="Frequency in 1000s", fill="Age Bin", title="Frequency of different Age Bins") +
  theme_minimal()

# Age Bucket wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Age_Bin, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="Age Buckets", y="Frequency in 1000s", fill="Performance Tag", title="Age Bucket wise Performance Tag Frequency")


```

### Gender

```{r}
#=============#
#   Gender    #
#=============#
# Summary for Gender
dem$Gender <- as.factor(dem$Gender)
dem$Gender %>%
  summary()

# 2 NA's

# Converting NA for Gender variable to "M"
levels(dem$Gender)[1] <- "M"

# Plot for frequency of each Gender
ggplot(dem, aes(x=Gender, y=..count../1000, fill=Gender)) +
  geom_bar() +
  scale_fill_manual(values = cp_2)+
  labs(x="Gender", y="Frequency in 1000s", fill="Gender", title="Frequency of different Gender") +
  theme_minimal()

# Gender wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Gender, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="Gender", y="Frequency in 1000s", fill="Performance Tag", title="Gender wise Performance Tag Frequency")


```

### Marital Status

```{r}
#=====================#
#   Marital Status   #
#=====================#

# Summary for Marital status at time of application
dem$Marital_Status %>%
  summary()
# 6 NA's

dem$Marital_Status <- as.factor(dem$Marital_Status)
# Converting NA for Marital status at time of application variable to "Married"
levels(dem$Marital_Status)[1] <- "Married"

# Plot for Marital status at time of application frquency
ggplot(dem, aes(x=Marital_Status, y=..count../1000, fill=Marital_Status)) +
  geom_bar()+
  scale_fill_manual(values = cp_8)+
  labs(x="Marital Status at time of application", y="Frequency in 1000s", fill="Marital Status", title="Frequency of different Marital Status") +
  theme_minimal()

# Marital Status wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Marital_Status, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="Marital Status", y="Frequency in 1000s", fill="Performance Tag", title="Marital Status wise Performance Tag Frequency")

```

### No of Dependents

```{r}
#=====================#
#   No of Dependents  #
#=====================#

# Checking for NA values
dem$No_Of_Dependents <- as.numeric(dem$No_Of_Dependents)
dem$No_Of_Dependents %>%
  is.na() %>%
  sum()

# 3 NA's

dem$No_Of_Dependents[which(is.na(dem$No_Of_Dependents))] <- 3

dem$No_Of_Dependents %>%
  as.factor() %>%
  summary()

# Checking for outliers
dem$No_Of_Dependents %>%
  quantile(seq(0,1,0.01), na.rm = T)

dem$No_Of_Dependents %>%
  boxplot(border = "#6fa058", outcol = "#ee853f")

#Converting the variable into factor type
dem$No_Of_Dependents <- dem$No_Of_Dependents %>% as.factor()

# Plot for No of Dependents Frequency
ggplot(dem, aes(x=as.factor(No_Of_Dependents), y=..count../1000, fill=as.factor(No_Of_Dependents))) +
  geom_bar() +
  scale_fill_manual(values=cp_5)+
  labs(x="No of Dependents", y="Frequency in 1000s", fill="No of Dependents", title="Frequency of No of Dependents") +
  theme_minimal()

# No of Dependents wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=No_Of_Dependents, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="No of Dependents", y="Frequency in 1000s", fill="Performance Tag", title="No of Dependents wise Performance Tag Frequency")



```

### Income

```{r}
#=============#
#   Income    #
#=============#

# checking for NA values
dem$Income %>%
  is.na() %>%
  sum()
# 0

# Checking for outliers
dem$Income %>%
  quantile(seq(0,1,0.01), na.rm = T)

dem$Income %>%
  boxplot(border = "#6fa058", outcol = "#ee853f")

dem$Income %>%
  as.factor() %>%
  levels()

# Converting Income less than 1 to 1.0
dem[(which(dem$Income < 1)), ] $Income <- 1.0

# Creating Income Bracket
# Income Bracket Function

income_bin <- function(income = 1){
  if(income >= 1 && income <=10)
    return ("1-10")
  else if(income >= 11 && income <=20)
    return ("11-20")
  else if(income >= 21 && income <=30)
    return ("21-30")
  else if(income >= 31 && income <=40)
    return ("31-40")
  else if(income >= 41 && income <=50)
    return ("41-50")
  else
    return ("51-60")
}


dem$Income_Bin <-  dem$Income %>%
  sapply(income_bin) %>%
  as.factor()

# Income Bucket wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Income_Bin, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="Income Buckets", y="Frequency in 1000s", fill="Performance Tag", title="Income Bucket wise Performance Tag Frequency")

```


### Education

```{r}
#===============#
#   Education   #
#===============#

# checking for NA values
dem$Education <- as.factor(dem$Education)
dem$Education %>%
  is.na() %>%
  sum()

# 0

# Checking for blank rows
dem$Education %>%
  summary()

levels(dem$Education)[1] <- "Professional"
attach(dem)
#Education <- as.factor(Education)
# Plot for Education Frequency
ggplot(dem, aes(x=Education, y=..count../1000, fill=Education)) +
  geom_bar() +
  scale_fill_manual(values=cp_5)+
  labs(x="Education", y="Frequency in 1000s", fill="Education", title="Frequency of Education") +
  theme_minimal()

# Education wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Education, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="Education", y="Frequency in 1000s", fill="Performance Tag", title="Education wise Performance Tag Frequency")

```


### Profession

```{r}
#=================#
#   Profession    #
#=================#

# checking for NA values
dem$Profession %>%
  is.na() %>%
  sum()

# 0

# Checking for blank rows
dem$Profession %>%
  summary()
dem$Profession <- as.factor(dem$Profession)
levels(dem$Profession)[1] <- "SAL"

# Plot for Profession Frequency
ggplot(dem, aes(x=Profession, y=..count../1000, fill=Profession)) +
  geom_bar() +
  scale_fill_manual(values=cp_3)+
  labs(x="Profession", y="Frequency in 1000s", fill="Profession", title="Frequency of Profession") +
  theme_minimal()

# Profession wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Profession, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="Profession", y="Frequency in 1000s", fill="Performance Tag", title="Profession wise Performance Tag Frequency")
```







### Type of Residence

```{r}
#=======================#
#   Type of residence   #
#=======================#
# checking for NA values
dem$Type_Of_Residence %>%
  is.na() %>%
  sum()

# 0

# Checking for blank rows
dem$Type_Of_Residence %>%
  summary()
dem$Type_Of_Residence <- as.factor(dem$Type_Of_Residence)
levels(dem$Type_Of_Residence)[1] <- "Rented"

# Plot for frequency of type of residence
ggplot(dem, aes(x=Type_Of_Residence, y=..count../1000, fill=Type_Of_Residence)) +
  geom_bar() +
  scale_fill_manual(values=cp_5)+
  labs(x="Type of residence", y="Frequency in 1000s", fill="Type of residence", title="Frequency of Type of residence") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=40, hjust=1))

# Type of Residence wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Type_Of_Residence, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = cp_2) +
  labs(x="Type of Residence", y="Frequency in 1000s", fill="Performance Tag", title="Type of Residence wise Performance Tag Frequency") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=40, hjust=1))

```

#### Observation 

* People who live in a rented apartment have a higher defaulter rate. This is also due to the fact that most of the customers are the ones who live in a rented apartment




### No of Months in Current Residence

```{r}
#===========================================#
#   Number of months in current residence   #
#===========================================#

# Checking for NA values
dem$Months_In_Current_Residence %>%
  is.na() %>%
  sum()

# 0

# Checking for outliers
dem$Months_In_Current_Residence %>%
  quantile(seq(0,1,0.01), na.rm = T)

dem$Months_In_Current_Residence %>%
  boxplot(border = "#6fa058", outcol = "#ee853f")


# Resident Years Bin Function
res_yrs_bin <- function(nom=0){
  noy = nom/12
  if(noy > 0 && noy < 1)
    return("< 1 yr")
  else if(noy >= 1 && noy < 2)
    return("1 yr")
  else if(noy >= 2 && noy < 3)
    return("2 yrs")
  else if(noy >= 3 && noy < 4)
    return("3 yrs")
  else if(noy >= 4 && noy < 5)
    return("4 yrs")
  else if(noy >= 5 && noy < 6)
    return("5 yrs")
  else if(noy >= 6 && noy < 7)
    return("6 yrs")
  else if(noy >= 7 && noy < 8)
    return("7 yrs")
  else if(noy >= 8 && noy < 9)
    return("8 yrs")
  else if(noy >= 9 && noy < 10)
    return("9 yrs")
  else
    return("> 10 yrs")
}

# Creating No of years in current residence variable
dem$Yrs_Curr_Res <- dem$Months_In_Current_Residence %>%
  sapply(res_yrs_bin) %>%
  as.factor()

# Plot of frequency of No of years in current residence variable
ggplot(dem, aes(x=Yrs_Curr_Res, y=..count../1000, fill=Yrs_Curr_Res)) +
  geom_bar() +
  labs(x="No of Years in residence", y="Frequency in 1000s", fill="No of Years in residence", title="Frequency of Years in residence") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=40, hjust=1))



# Years In Current Residence wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Yrs_Curr_Res, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = cp_2) +
  labs(x="Years In Current Residence", y="Frequency in 1000s", fill="Performance Tag", title="Years In Current Residence wise Performance") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=40, hjust=1))


```


#### Observation

* Most of the customers are the ones who live in a place they moved in within the last year. Therefore, most number of defaulters are from this group. 


### Number of Months in Current Company

```{r}
#=========================================#
#   Number of months in current company   #
#=========================================#

# Checking for NA values
dem$Months_In_Current_Company %>%
  is.na() %>%
  sum()

# 0

# Checking for outliers
dem$Months_In_Current_Company %>%
  quantile(seq(0,1,0.01), na.rm = T)


dem$Months_In_Current_Company %>%
  boxplot(border = "#6fa058", outcol = "#ee853f")

# Capping No of months in current company to 74
dem[(which(dem$Months_In_Current_Company > 74)),] $Months_In_Current_Company <- 74

#   Current Company Years Bin Function
comp_yrs_bin <- function(nom=0){
  noy = nom/12
  if(noy > 0 && noy < 1)
    return("< 1 yr")
  else if(noy >= 1 && noy < 2)
    return("1 yr")
  else if(noy >= 2 && noy < 3)
    return("2 yrs")
  else if(noy >= 3 && noy < 4)
    return("3 yrs") 
  else if(noy >= 4 && noy < 5)
    return("4 yrs")
  else if(noy >= 5 && noy < 6)
    return("5 yrs")
  else
    return("> 6 yrs")
}

# Crating variable No of years in curr comp
dem$Yrs_Curr_Comp <- dem$Months_In_Current_Company %>%
  sapply(comp_yrs_bin) %>%
  as.factor()

# Plot for No of years in current company
ggplot(dem, aes(x=Yrs_Curr_Comp, y=..count../1000, fill=Yrs_Curr_Comp)) +
  geom_bar() +
  labs(x="No of Years in Current Company", y="Frequency in 1000s", fill="No of Years in Current Company", title="Frequency of Years in Current Company") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=40, hjust=1))


# Years In Current Company wise Performance Tag Frequency
dem %>%
  filter(!is.na(Performance_Tag)) %>%
  ggplot(aes(x=Yrs_Curr_Comp, y=..count../1000, fill=as.factor(Performance_Tag))) +
  geom_bar(position = "dodge") +
  theme_minimal()+
  scale_fill_manual(values = cp_2) +
  labs(x="Years In Current Company", y="Frequency in 1000s", fill="Performance Tag", title="Years In Current Company wise Performance Tag Frequency")


```

### Data Split

```{r}

dem = subset(dem, select = -c(Age, Income, Months_In_Current_Company, Months_In_Current_Residence) )
set.seed(123) 
split = sample.split(dem$Performance_Tag, SplitRatio = 0.70) 
  
training_set = subset(dem, split == TRUE) 
test_set = subset(dem, split == FALSE) 

```

```{r}

```

```{r}
# Plot data
#dem$Performance_Tag <- as.factor(dem$Performance_Tag)
ggplot(data = dem, aes(x = Marital_Status+No_Of_Dependents+Education+Profession+Type_Of_Residence+Age_Bin+Yrs_Curr_Res+Income_Bin+ Yrs_Curr_Comp, y = Performance_Tag, color = as.factor(Performance_Tag), shape = as.factor(Performance_Tag))) + 
  geom_point(size = 2) +
  scale_color_manual(values=c("#000000", "#FF0000")) +
  theme(legend.position = "none")
```


### Fitting SVm to the training set

```{r}
classifier = svm(formula = Performance_Tag ~  Marital_Status+No_Of_Dependents+Education+Profession+Type_Of_Residence+Age_Bin+Yrs_Curr_Res+Income_Bin+ Yrs_Curr_Comp,
                 data = training_set, 
                 type = 'C-classification', 
                 kernel = 'radial',
                 gamma=0.5) 

```

```{r}
print(classifier)
```

```{r}
y_pred = predict(classifier, newdata = test_set[-8]) 

length(y_pred)
length(test_set$Performance_Tag)

```

```{r}
cm = table(test_set$Performance_Tag, y_pred) 
cm
```

